
% Strong background references addresses "what is already known" 
The idea of developing climbing robots was expected for years but only in 1986, 
Nishi et al. presented the design of a robot that can walk on a vertical surface
in their seminal paper \cite{Nishi1986}. During the last thirty years, many prototypes
of climbing robots were proposed for specific applications to replace humans 
in dangerous and difficult tasks including, inspection and maintenance \cite{Song2008}, 
cleaning tall structures \cite{Nansai2016}, search and rescue missions \cite{Eich2008}. 
However, % Mention a Gap in knowledge
there are still other intelligent ways to climb walls not yet studied or reported 
\cite{bio_inspired2015}. %Introduction of the problem 

The research work done on climbing robots is reviewed and classified in 
\cite{Fang2022} and \cite{A.Hajeeretall2020} respectively, according to the adhesion 
method \cite{longo2008}, used to attach the robot to the wall and locomotion \cite{Chu2010}, 
type of the walking mechanism. These are considered as the two main problems to be solved in 
designing wall-climbing robots. Other requirements may also be requested depending on the 
desired task, namely: R1) to move safely, and R2) fast in emergency situations, R3) ability 
to carry a payload, and R4) avoiding obstacles, .. etc. 
%Providing our solution for first problem: attachment technique 

To solve the mentioned challenges. Firstly, we propose in this paper using a rope as 
an adhesion mechanism for the following reasons: 
% Advantages of using a rope : limits wiht high inclinatoions
It is well known that there is a limit in the inclination that a mobile robot can 
address related to the value of the friction coefficient at the contact. Indeed, 
in a slope with a very high inclinations \cite{Abdalla12020}, the robot eventually 
creates a tangential force on the terrain that surpasses the friction force that is 
needed to prevent slippage and falling. 

%Advantages of a rope
Instead, the use of a rope will allow the contact forces to better satisfy the 
friction constraints (i.e., lay more in the middle of the friction cones) when 
walking on highly inclined terrains. Indeed, a rope introduces an external force 
to compensate for gravity solving this issue and allowing the robot to walk on 
ideally any terrain inclination (even vertical).

Additionally, in the case of legged robots with a rope can help to keep joints 
far from the kinematic limits in a more kinematically advantageous configuration 
without the need to lean back or forth to keep stability. In other words, the rope 
will ensure inherent safety (R1) from falling and efficiency when rappelling down 
(i.e. exploiting the effect of gravity to unwind the rope.)  

%Advantages of using a rope
Furthermore, with respect to most common climbing animals that uses sticky pads 
(e.g., geckos) and gaits to climb up/down a wall \cite{Kim2008}, \cite{Riskin2009}, 
the aid of a rope as the recently bio-inspired dragline locomotion from spiders in 
\cite{Wang2014} can dramatically increase the locomotion speed (R2) by a 
winding/releasing mechanism, therefore being a preferable solution in applications 
that require a prompt intervention such as: rescue missions. 

%Last advantage of rope
Finally, the amount of payload that can be carried by sticky/adhesive based climbing 
robots is very limited due to the limits in the tangential component of the adhesive 
and the leg actuation. While, a feature of rope being inextensible (on a certain extent) 
will allow the amount of payload (R3) that the robot is able to carry to be much higher 
and it is only limited by the maximum force deliverable by the winding mechanism (e.g. 
an hoist), which using a gearbox can be designed to be very high. 

%Difficulties due to the dynamics when using a rope
After all, having the robot attached to a rope poses many challenges: one is the 
under-actuation (the robot is not able to control the position of its center of mass 
while not in contact) and the other is that a rope represents  a unilateral constraint 
(can only pull and not push), which further complicates the already hybrid dynamics and 
the low control authority of this class of robots. 

%Providing a solution for second problem: locomotion technique, we can reach the target with jumps 
Secondly, we also suggest using the jumping as a moving method to reach a desired location 
on te wall. Indeed, rather than slowly taking steps to move, the robot can take one or multiple 
jumps to reach the target on the wall. For each jump,  the winding/releasing mechanism can be 
exploited to behave like a variable length pendulum (see Fig. \ref{fig:sp}). The jumping 
locomotion is similar to the Salto-1P jumping robot shown in \cite{Haldane2017} but the key 
feature of jumping with rope relies in the ability to address terrains of high inclination 
(up to vertical). 

%Difficulties in the jumping technique
When taking a jump with a rope that can be released, the resulting motion of the robot (and 
so the possibility to reach the target) will depend on two factors: 1) the initial impulse 
that the robot exerts on the wall at lift-off, 2) the time-law the rope is released/re-winded.
%Optimization for plannig 
Therefore, a planning strategy for these kind of motions, to be successful, it should take 
into account both these two factors together with the under-actuation of the robot and the 
physical constraints, namely: the constrains posed by the rope, the actuator limits and the 
contact interaction (i.e. friction).

Numerical optimization is an attractive solution for this planning problems 
\cite{Nguyen2019,Ding2020} because it allows to jointly cope with these aspects in an optimal 
fashion while ensuring that the constraints are satisfied in a planning horizon resulting in a 
feasible trajectory. Casting the locomotion planning as an optimization problem allows one to 
represent high level tasks and system dynamics using cost functions and constraints. 

In this framework different goals can also be pursued. for instance, minimize energy consumption 
or hoist effort, conversely, the jump time can be also minimized to achieve the target in a 
shortest time in case of emergencies or rescue missions and extend mission duration. 


\subsection{Proposed Approach and Contribution}
In this work, we present a novel robotic platform called CLIO that is able to 
reach desired targets on a vertical wall. We propose planning approaches 
based on numerical optimization to solve the jumping problem employing a simplified model of 
the dynamics.
To summarize, the contributions of the paper are:
\begin{itemize}
	\item a novel design of a jumping robot platform CLIO 
	\item a planning approach, based on optimal control, to generate a desired jump motion, 
	overcoming an obstacle comparing pros and cons of each approach.
	\item simulations and experiments to validate  the effectiveness of the proposed approach 
	in a simulation with a simplified model and in a realistic (gazebo) simulation with the 
	full dynamics of the robot.
\end{itemize}


\subsection{Outline}

The paper is organized as follows: Section II gives an
overview of the robot ....
on III describes the optimization
problem with the LIP model and how it is used to compute
CoM position velocity and contact force references. Simula-
tions and experiments with Aliengo robot are illustrated in
Section IV. Finally, we draw the conclusions in Section V.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
