% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[letterpaper, 10 pt, conference]{ieeeconf}


\overrideIEEEmargins                                      % Needed to meet printer requirements.

\IEEEoverridecommandlockouts                              % This command is only needed if 
% you want to use the \thanks command

% include the list of acronyms, math commands and new commands used in this paper
\input{includes/used_packages.tex}
\input{includes/acronyms.tex}
\input{includes/math_commands.tex}
\input{includes/newcommands.tex}


\title{\LARGE \bf CLIO: a Novel Robotic Solution for Exploration and Rescue Missions in Hostile Mountain Environments}
\author{Michele Focchi$^{1}$, Mohamed Bensaadallah$^{2}$, Marco Frego$^{3}$, Angelika Peer$^{3}$, \\Daniele Fontanelli$^{4}$, Andrea Del Prete$^{4}$, Luigi Palopoli$^{1}$\vspace{-1.0cm}
\thanks{$^1$ The authors are with the Dipartimento di Ingegneria and Scienza dell'Informazione (DISI), University of Trento. Email:  \href{mailto:name.surname@unitn.it}{name.surname@unitn.it}}
\thanks{$^2$ The author is with the Department of Electronics, University of Batna 2, Algeria. Email: \href{mailto:m.bensaadallah@univ-batna2.dz}{m.bensaadallah@univ-batna2.dz}}
\thanks{$^3$ The authors are with the Faculty of Science and Technology, Free University of Bozen-Bolzano. Email:  \href{mailto:name.surname@unibz.it}{name.surname@unibz.it}}
\thanks{$^4$ The authors are with Dipartimento di Ingegneria Industriale (DII), University of Trento Email:  \href{mailto:name.surname@unitn.it}{name.surname@unitn.it}. The publication was created with the co-financing of the European Union FSE-REACT-EU, PON Research and Innovation 2014-2020 DM1062 / 2021.} }


\begin{document}
\maketitle
\thispagestyle{empty}
\pagestyle{empty}


\begin{abstract}%150-250 word abstract
  Rescue missions in mountain environments are hardly achievable by
  standard legged robots---because of the high slopes---or by flying
  robots---because of limited payload capacity.  We present a 
  concept for a rope-aided climbing robot which can negotiate
  up-to-vertical slopes and carry heavy payloads.  The robot is
  attached to the mountain through a rope, and it is equipped with a leg
  to push against the mountain and initiate jumping maneuvers.
  Between jumps, a hoist is used to wind/unwind the rope to move
  vertically and affect the lateral motion.  This simple (yet
  effective) two-fold actuation allows the system to achieve high
  safety and energy efficiency. Indeed, the rope prevents the robot
  from falling while compensating for most of its weight, drastically
  reducing the effort required by the leg actuator. We also present an
  optimal control strategy to generate point-to-point trajectories
  overcoming an obstacle.  We achieve fast computation time ($<$1 s)
  thanks to the use of a custom simplified robot model.  We validated
  the generated optimal movements in Gazebo simulations with a
  complete robot model with a $<5\%$ error on a 16 $m$ long jump, 
  showing the effectiveness of the proposed
  approach, and confirming the interest of our concept.  Finally, we
  performed a reachability analysis showing that the region of
  achievable targets is strongly affected by the friction properties
  of the foot-wall contact.
\end{abstract}


\section{Introduction}\label{sec:introduction}
\input{sections/intro}

\section{Robot Description}\label{sec:robot}
\input{sections/robot}


\section{Motion Planning}\label{sec:motionP}
\input{sections/motion}


%\section{Motion Control} \label{sec:motionC}
%\input{sections/control}


\section{Simulation Results}\label{sec:results}
\input{sections/results}





\section{Conclusions}
\label{sec:conclusion}
\input{sections/conclusions}
\small
\bibliographystyle{IEEEtran}
\bibliography{references/references}

%\section{Acknowledgements}
%The publication was created with the co-financing of the European Union FSE-REACT-EU, PON Research and Innovation 2014-2020 DM1062 / 2021.

\end{document}

